### 2014-12-09 prep_figures/commute/curvature-MCT.Rmd
new file:   curvature-MCT.Rmd

---
title: "Ricci and commute time"
runtime: shiny
output: html_document
---

```{r, echo=FALSE}

library(ggvis)
library(ape)
library(apTreeshape)

source("process.R")

d <- get_results("../../results-rspr/commute6.mat")
d$dist <- as.factor(d$dist)

d %>%
  ggvis(x = ~kappa, y = ~MCT) %>%
  layer_points(size = 5, fill = ~dist) %>%
  # A dummy layer so that we can use text in the tooltip.
  layer_text(text := ~info, fontSize := 0) %>%
  add_axis("x", grid = FALSE) %>%
  add_axis("y", grid = FALSE) %>%
  add_legend(c("fill"), properties=legend_props(legend=list(y=0))) %>%
  add_tooltip(function(df) df$info)
```