### 2014-12-18 prep_figures/commute/commute-detail.Rmd

---
title: "Ricci and transit times"
runtime: shiny
output: html_document
---

```{r, echo=FALSE}

library(ggvis)
library(ape)
library(apTreeshape)

source("process.R")

# d <- get_results("../../results-rspr/commute-nodata/commute6.mat")
# d$dist <- as.factor(d$dist)

d5 <- get_results("../../results-rspr/commute-onesplit/123_45-1v5c.tsv")
d6 <- get_results("../../results-rspr/commute-onesplit/123_456-1v5c.tsv")
d <- d5

d %>%
  ggvis(x = ~kappa, y = ~MAT) %>%
  #ggvis(x = ~kappa, y = ~MCT) %>%
  layer_points() %>%
  # A dummy layer so that we can use text in the tooltip.
  layer_text(text := ~info, fontSize := 0) %>%
  add_axis("x", grid = FALSE) %>%
  add_axis("y", grid = FALSE) %>%
  add_tooltip(function(df) df$info)
```