### 2014-12-18 prep_figures/commute/commute-detail.Rmd
commute time plots

---
title: "Transit time distributions."
runtime: knitr
output: html_document
---

```{r, echo=FALSE}

library(ggplot2)

grab_data <- function(fname) {
  df <- read.table(fname, stringsAsFactors = FALSE)
  colnames(df) <- c("t1", "t2", "time", "count")
  df$cmp <- paste(df$t1, df$t2)
  df$dens <- df$count / sum(df$count)
  df
}

d <- rbind(
  grab_data("runs/run1-uniq_trees_T-col-CCC1_2D_3D_C4_5DD-CC1_C2_3DD_C4_5DD.tab"),
  grab_data("runs/run1-uniq_trees_T-col-CCCC1_2D_3D_4D_5D-CCCC1_2D_3D_5D_4D.tab"),
  grab_data("runs/run1-uniq_trees_T-col-CCCC1_2D_3D_4D_5D-CCC1_2D_3D_C4_5DD.tab"))

sub_d <- subset(d, t1 != t2)

ggplot(sub_d, aes(x=time, y=dens, color=cmp)) + geom_line() + scale_x_continuous(limits=c(0,100))
```