### 2014-12-19 prep_figures/commute/access-distn.Rmd
Less bad names for access distribution stuff.

---
title: "Transit time distributions."
runtime: knitr
output: html_document
---

```{r, echo=FALSE}

library(ggplot2)

grab_data <- function(fname) {
  df <- read.table(fname, stringsAsFactors = FALSE)
  colnames(df) <- c("t1", "t2", "time", "count")
  df$cmp <- paste(df$t1, df$t2)
  df$dens <- df$count / sum(df$count)
  df
}

data_of_num <- function(n) {
  df <- grab_data(paste0("runs5/run1.",as.character(n),".tab"))
  df$num <- as.factor(n)
  df
}

d <- do.call(rbind, lapply(c(7,14,66,68,69), data_of_num))

sub_d <- subset(d, t1 != t2)
sub_d <- subset(d, t1 < t2)

ggplot(sub_d, aes(x=time, y=dens, color=num)) + geom_line() + scale_x_continuous(limits=c(0,25))
```