### 2015-02-19 AlexsNotes.tex
Trivial notes on ranked NNI.

\documentclass[12pt]{article}

\usepackage{amsmath,amsthm}
\usepackage{todonotes}
\usepackage[notref,notcite]{showkeys} % show keys for eqs, etc.
\usepackage{cite}
\usepackage{enumerate}

\newtheorem{lemma}{Lemma}

%\newcommand{\deg}{\mathrm{deg}}

\begin{document}

We consider the ranked-NNI graph (the distance on ranked topologies inherited from $\tau$-space).

\begin{lemma}
Let $T$ be a ranked tree on $n>2$ leaves. Then $$n-1\leq \deg(T)\leq2(n-2).$$ 
\end{lemma}

\proof
Obvious. 
\endproof

\begin{lemma}
\begin{enumerate}[(1)]
\item $\frac{\deg(T)}{deg(R)}>\frac12.$
\item $deg(T)-deg(R)> n-3.$
\end{enumerate}
\end{lemma}

\proof
Follows from previous Lemma. 
\endproof

\begin{lemma}
$d_{\tau}(T,R) = 1 \Rightarrow |N_1(T)\cap N_1(R)|\in\{0,1\}.$
\end{lemma}

\proof
Obvious.
\endproof

\end{document}