### 2015-02-19 AlexsNotes.tex
Lower bound for curvature and an abstract question.

\documentclass[12pt]{article}

\usepackage{amsmath,amsthm}
\usepackage{todonotes}
\usepackage[notref,notcite]{showkeys} % show keys for eqs, etc.
\usepackage{cite}
\usepackage{enumerate}

\newtheorem{lemma}{Lemma}
\newtheorem{question}{Question}

%\newcommand{\deg}{\mathrm{deg}}

\begin{document}

We consider the ranked-NNI graph (the distance on ranked topologies inherited from $\tau$-space).

\begin{lemma}
Let $T$ be a ranked tree on $n>2$ leaves. Then \[n-1\leq \deg(T)\leq2(n-2).\] 
\end{lemma}

\proof
Obvious. 
\endproof

\begin{lemma}
\begin{enumerate}[(1)]
\item $\frac{\deg(T)}{deg(R)}>\frac12.$
\item $deg(T)-deg(R)> n-3.$
\end{enumerate}
\end{lemma}

\proof
Follows from previous Lemma. 
\endproof

\begin{lemma}
$d_{\tau}(T,R) = 1 \Rightarrow |N_1(T)\cap N_1(R)|\in\{0,1\}.$
\end{lemma}

\proof
Obvious.
\endproof

\begin{lemma}
Assume uniform random walk and $d_\tau(T,R) = 1$. Then \[\kappa(T,R) \leq \frac{1}{2(n-2)}.\]
\end{lemma}

\proof
Follows from Lemma~5.2 in WhiddenMatsen. 
\endproof

\begin{question}
Assume some (e.g.\ uniform, lazy, ...) random walk. What are the $m$ and $M$ such that 
\[
m(\deg x,\deg y, \ldots) \leq \kappa(x,y) \leq M(\deg x, \deg y, \ldots)? 
\]
\end{question}

\end{document}