### 2016-11-29 response_to_reviewers.txt
(most) revisions up to page 20

Reviewer 2

Page 1

27: The spelling 'multimodal' seems to be more common than 'multi-modal' (and would be in line with other uses in your paper; e.g. page 3, line 12).
* corrected

30: The word 'the' seems to be missing before 'study'.
* corrected

32: The word 'the' seems to be missing before 'Ricci-Ollivier'.
* corrected

36: Here, I would suggest clarifying the phrase 'and n their number of leaves' a bit. In its current state, it might not be clear from the wording whether each tree has n leaves or whether n is the total number of leaves of all trees.
* revised to "... and n is the number of leaves in each tree, ..."

44-46: The phrasing 'We show ... showing ...' sounds a bit odd because 'show' is used twice.
* revised "showing" to "demonstrating"

Page 2
10: It looks like some kind of connecting phrase is missing between 'particular' and 'properties' (e.g.  'with respect to').
* we added the suggested connecting phrase

11: The word 'the' seems to be missing before 'investigation'.
* corrected

52-53: As you start this part of this sentence with 'both', an 'and' seems to be more fitting in 'leaves [12, 13] or for SPR'.
* corrected

54: Do you rather mean 'an infinite number of leaves' or 'infinitely many leaves' instead of 'infinite leaves'?
* corrected to "infinitely many leaves"

Page 3

15: A more fitting position of 'still' seems to be between 'is' and 'known'.
* corrected

18: A word like 'the' or 'its' seems to be missing before 'degree'.
* corrected to "its".

30: The phrase 'begun to allow' sounds a bit odd. Swapping the verbs ('allowed to begin') might be preferable.
* The phrase "allowed to begin" would change our intended meaning, so we instead changed this to "... efficient fixed-parameter algorithms have recently been developed [17, 24, 25] that begin to allow such investigations." .

31: It looks like the word 'investigation' is missing the indefinite article or should be in the plural form.
* corrected, as noted above

33: The word 'the' seems to be missing before 'Ricci'.
* corrected

47: The word 'the' seems to be missing before 'curvature'.
* corrected

56: The same problem as on page 1, line 36.
* corrected to "... each of which has n leaves".

Page 4
44: In this context, 'linear-time' seems to be more common without a hyphen.
* corrected

47-52: Both sentences use the phrase 'we found'. Varying the wording a bit would make it sound better.
* We revised the second sentence with "we found" to begin with "Our results demonstrated that ...".

48: The word 'the' seems to be missing before 'distance'.
* corrected

Page 5

27: The word 'the' seems to be missing before 'proba-'.
* corrected

29: The word 'the' seems to be missing before 'knowledge'.
* corrected

Page 6

30: Is it really necessary to define the LCA of two or more subtrees as the LCA of their parent nodes (and not as the LCA of the root nodes of the subtrees)? It seems a bit unconventional (also w.r.t. your LCA definition of one or more nodes beforehand).
* Corrected. We did indeed mean the root nodes of the subtrees by "parent".

40: Either the word 'the' seems to be missing before 'arbitrary' or it has to be 'arbitrarily' (in which case the 'of' before 'trees' has to be removed).
* corrected by adding "the".

Page 7

13: The wording 'we will specialize to that space being a graph' sounds odd. Maybe something along the lines of 'we restrict ourselves to ...' is preferable.
* revised to "... we restrict ourselves to graphs equipped with a shortest-path distance."

20: It looks like the apostrophe in 'movers' is missing (it should be 'mover's').
* corrected

Page 8

14: It looks like there is too much white space behind 'e.g.'.
* corrected

24: In this context, 'previously' seems more appropriate than 'previous'.
* corrected

28-33: Both points are possibly related.

1. In the final sentence of the first paragraph, m is assumed to be (2n-3)!! ('m is the rapidly growing function (2n-3)!!'). You continue using m in the second paragraph. Is it an arbitrary value not larger than (2n-3)!! (or still equal to the maximum)? I would suggest clarifying the transition.
* As stated in the second sentence of the first paragraph, $m$ is the number of trees in the graph. This is (2n-3)!! for the full rSPR graph, as noted in that final sentence. In the second paragraph, $m$ still refers to the number of trees in the graphs in question, which is upper bounded by (2n-3)!! as you note. We have revised these sentences to limit confusion and address your next point. We now note that, for the full rSPR graph, the value of $m$ is given by $(2n-3)!! in the first paragraph. We removed the redundant reference to $m$ in the first sentence of the second paragraph.

2. The phrase 'graphs from subsets of m high probability trees' is a bit ambiguous. Does it mean that the graphs are constructed from subsets which themselves consist of m trees or are the graphs constructed from subsets of the m-element set of high probability trees?
* The subsets consist of m trees. As noted above, m is simply the number of trees in a given graph and not a fixed value. We removed the redundant reference to m here.

47-48: What is the unit of the length of the Newick representation?
* The unit of the length is the number of characters in the string (equivalently wrt big O, the size in bytes). We realized in a follow up paper that, although the length of a Newick string is typically assumed to be linear, this does require some (reasonable) assumptions. We now indicate so here and cite the preprint of that paper.

51: To which list does 'in this list' refer to?
* This statement refers to the order of the tree in the specified ordering. We revised "in this list" to "in this ordering" to prevent confusion.

Page 9
44-45: The statement starts with 'requires only O(n)-time for each tree' (of which there are m).  Therefore, it is not quite clear (even though not wrong), why the total time of this step is bound by the unnecessarily unsharp term 'O(mn^3)'.
* This is O(n)-time for each neighbor tree, of which there are O(mn^2), for a total of O(mn^3)-time. We added the word neighbor to clarify.

Page 10

42: The wording 'within a subtree' seems a bit imprecise and could be clarified by changing it to something along the lines of 'within the subtree to be moved / regrafted'.
* revised to "within the subtree to be moved".

55: The term 'every' seems to be more fitting than 'each' in this context.
* corrected

Page 11

9-12: The wording seems a bit imprecise because the target of a move is not the node itself but the node's edge.
* The wording was indeed confusing. We are assigning moves for counting
	purposes to one possible source node, not specifying a move to any
	given node. We added the sentence "To do so, we assign each of the
	neighbors to one node of the tree, such that any given neighbor can be
	obtained by moving the subtree rooted at its assigned node." We then
	changed each mention of assigning moves to assigning neighbors.

23-25: The case distinction requires that the extra root node ρ has depth -1, the actual root depth 0 and so on, but this is stated nowhere in the paper.
* We added a sentence defining the depth of the rho node to be -1 to the
	preliminaries: "The depth of \rho is considered to be -1, the depth of
	the original root is 0, and the depth of each other node is given by
	its distance from the original root.

33: Probably, the set-difference operator ('\') is supposed to denote the pruning (including the merging of the incoming and outgoing edge of the resulting unary node). Unfortunately, this is not explained. A short explanation would make this more obvious to the reader.
* We have added an explanation here: "This implies that both T \U = S \U
	and T \V = S \V by the definition of an rSPR operation, where \ here
	indicates removing the subtree and its parent edge and then replacing
	its former parent node and both adjacent edges with a single edge."

Page 12

15: It looks like that '-th' is missing behind the '|r|' in order to turn it into an ordinal number.
* corrected

23: What is the unit of the 'O(n) space'?
* bits of memory, as is standard for the Big-O notation with respect to space.

52-55: The significance of this paragraph could be increased by stating the special features, properties or outcomes of these runs instead of barely stating that you have performed them.
* The runs are analyzed in the next section. To avoid confusion we moved
	the description of the runs to the Access time simulation section and
	simply say "We used this procedure to sample random walks on varying
	size rSPR graphs for our analysis in the next section" here.

Page 13
32: Which graph is 'the rSPR graph ... as discussed in Section 3.1'? The complete one with (2n-3)!!  trees? This is not quite clear since section 3.1 also considers other values for m.
* This refers to the complete graph. We revised our use of m as
	described above, to limit this type of confusion. In addition we added
	the word full ("full rSPR graph") to this sentence.

33: It is stated that the curvature is computed for 'given pairs of trees', but who / what gives these pairs? The wording is a bit odd here.
* This is explained in the next paragraph. We could not compute
	curvatures for all pairs of trees directly, but instead compute
	curvatures for each tanglegram. The procedure described here is to
	used to compute the curvature of any given pair of trees.

55: The word 'tempered' is slightly ambiguous, because it can have the meaning 'to moderate' as well as 'to harden'.
* revised to "reduced"

****TODO FIGURES START*****

Page 14

56-: The connection between a smaller κ and a right-shift of the access time distribution is difficult to see as differences in the saturation are hardly noticable.

Page 15

Plots of Figure 5:
1. It is very hard to extract any information from the plots, because the plots (as soon as there are a few more lines) quickly get congested due to overlaps and hardly (or not at all) visible differences in saturation.

2. Adding axis labels (and not just explaining them in the figure caption) could help the readability of the plots.

3. It might be a bit counterintuitive to associate smaller κ values with an increased saturation.

4. Furthermore, the print version of articles in this journal usually come with monochrome images, (then) making it basically impossible to distinguish the lines by colour and saturation.

Page 16
Plots of Figure 6:
1. Except for the axis labels, the plots have the similar problems as those of Figure 5.
2. Plots (c) and (d) are especially problematic w.r.t. overlaps.
3. Adding brightness (blue vs. light blue) as another distinctive feature (besides saturation) can
also hamper the readability (when the differences are as small as they are here).

****TODO FIGURES END*****

45-56:

1. Which cases (of the four cases (a) – (d)) are meant with 'In both cases'?
* This referred to both of the cases mentioned in text (early and later
	access times). We revised the phrase to "these plots" to avoid
	confusion.

2. What is the rationale (and meaning) behind '[excluding] pairs that are simple inclusions of pairs of trees with a smaller number of taxa'?
* These plots demonstrate the effect of curvature on access times when variables such as degree are controlled. Tree pairs that are inclusions of pairs of trees with a smaller number of taxa have largely unchanged access times to the smaller trees and clutter the figure.
	A pair of trees that both contain leaves 1 and 2 as siblings are an
	example of inclusions of the pair of trees obtained by merging those siblings into a single leaf and relabeling the leaves.

Page 17

12-15: Adding hyphens (e.g. 'distance-1 pairs') might increase the readability.
* We have made the suggested change

15-23: The traceability / comprehensibility of the statements would strongly benefit from improving the readability of the plots in Figure 5 and 6.
* TODO

53: It looks like there is too much white space behind 'w.l.o.g.'.
* corrected

Page 18

Proof of (i):The proof only works when the '+14' in the numerator is not reduced to '+12' (which is also not necessary, because bounding the sum of the logarithms in the denominator is sufficient.
* TODO: Do you see a problem here Erick? Reducing the numerator from 14
	to 12 is valid. Maybe he is confused by the last step, where I borrow
	from the reduced -n in the denominator to reduce the 20 to 18?

Proof of (ii):
There seems to be a small error in the format. The penultimate equality is not aligned with the other equal and less-than signs.
* corrected

42-43: The statement 'The maximum degree difference from Lemma 5.2. occurs between trees that differ by many rSPR operations.' is (in this generality) not correct or at least misleading. Consider the rSPR graph for X-trees with four leaves from Figure 3(a). The ladder tree and the balanced tree have the maximum degree difference but are only separated by a single rSPR move while there are other tree pairs separated by two moves (which, in this context / example, should be considered as many).
* revised to "may occur". Our intent here is simply to lead up to our
	improved bounds for adjacent trees.

Page 19

39:
1. There seems to be an inconsistency between the definition of internal edges and their number stated in this line. On page 5, internal edges are defined as those that do not contact a leaf or ρ. Then, a 'rooted tree with k leaves' cannot have 2(k-1) internal edges as stated here (because this is the number of all edges (also those leading to leaves), disregarding the extra root edge). So, either the definition of 'internal edge' has to be adopted or the word 'internal' has to be dropped here.
* We have dropped the word "internal" here.

2. The white space after 'edges' is missing.
* corrected

Page 20

9-19: While Corollary 5.4 appears correct, the given justification seems to be a bit problematic.  Based on the definition of γ and the excluded (subtracted) moves, it appears that these γ moves still include some which move R. For example, from the set of all possible moves for T (deg(T)), only some moves into R (2kb, case (ii)) and some moves of R's parent subtree (2(j-1), case (iv)) are excluded.  However, it is then stated that these moves 'can be applied to T and S with the same source and target nodes'. The minor (but arguable) problem concerns the part 'same source nodes': When are they the same? Is the parent node of R the same although its position differs in T and S? If not, then not all γ moves can be applied with the same source nodes. The bigger problem is created by the fact that moves moving R with the same target node eliminate the only difference between T and S and, thus, the difference between the obtained neighbours is zero (not one).  Overall, it seems to be a bit more complicated to describe how the desired (T',S') pairs can be obtained.
* revised the first part with "We consider the parent node of R in both T and S to be the same source node for this purpose.". 
* TODO: We can either change the distance bound here to <= 0 or subtract
	out the distance 0 pairs. Check which is better after completing the
	other revisions.

14: The phrase 'the pairings are disjoint' might need some more explanation as 'disjoint' might get interpreted in several ways (e.g. no tree is allowed to occur in more than one pairing vs. the pairings are pairwise distinct).
* revised to note "...  each neighbor of T is paired
with a single neighbor of S, and vice versa. Note that a tree may be a
neighbor of both T and S and thus involved in two different pairings."

37: The word 'and' seems to be missing before 'then' .
* corrected

Lemma 5.5:
1. Concerning the proof of (i): it might be helpful to elaborate on why a case distinction is
necessary (Why is it necessary to increase j? Why are the subcases of (not) moving the additional leaf necessary?) in order to clarify the line of argument.
* TODO

2. Concerning statement (ii): Is there a particular reason why the quotient is bounded for n ≥ 4?
The overall lemma assumes n ≥ 3, which would also lead to a nice bound (0.8).
*

Page 21
49:
1. The meaning of the function L(.) can be guessed from the context without too much effort,
but providing a short definition / explanation would help avoiding any misconceptions.
2. Does this equality really hold between the restricted versions of T and S , which differ only in
R and, thus, should still be different after restricting them to X \ L(Q)?
51-53: Using U as a subtree while it also denotes a whole tree in the neighbourhood of T and S just
a few lines before can be slightly confusing.
Page 22
30-32: The statement on the proportionality of small moves and small subtrees seems to be a bit
troubled.
1. For example, if both children of the root are not small, each small subtree with one leaf
contributes 2n-6 moves, while each small subtree with two leaves contributes 2n-8 moves.
So, what is the factor of proportionality?
2. In the phrase 'where one of the root's children is small', is the intended meaning exactly one
or at least one? Assuming, e.g., 2n-6 as the factor of proportionality (coming from (the
majority of) the leaves), the statement does not hold in general (neither for exactly nor for at
least). Consider, e.g., a tree with a leaf as the only small child of the root and at least three
small subtrees with two leaves (for the case exactly) respectively the balanced tree with four
leaves (for the case at least).
36: Writing 'one 2-leaf subtree' instead of '1 2-leaf subtree' appears more readable (and is more
consistent with the previous use of numerals).
38-42:
1. Since Lemma 5.7 shall hold for n ≥ 3, the stated number of small moves (2n2-4n-6) is not
correct for all intended n. The reason is that for the ladder tree with three leaves, the small
subtree with two leaves is also a child of the root and, thus, contributes 2n-3-3 (not 2n-3-5)
moves. Hence, a case distinction seems to be necessary here and for the inequality in the
subsequent paragraph (lines 40-42)
2. The chain of (in)equalities in lines 40-42 does not run through, because adding (24-8n) leads
to the n-free term '+12' (not '+16'). But this can be 'repaired' through the necessary case
distinction n = 3 vs. n ≥ 4).
45+55: Stating the explanation of the size '(in terms of the number of leaves)' directly in line 45
possibly avoids some confusion (which can occur when reading Lemma 5.8 while having the
typical notion of size in mind).
54-55: The statement 'a ladder tree has the greatest average moved subtree size' does not seem to
hold for all n ≥ 4. For n = 4, the ladder tree has an average moved subtree size of 1, while it is 4/3
for the balanced tree. Similarly, the balanced tree has a larger one for n = 8 (1.55 vs. 1.39).
This might pose a problem to the proof because it based on the idea that the ladder tree leads to themaximum.
56: Similar to line 36, 'one leaf' and 'one subtree' appear more readable / consistent.
Page 23
11-13: The vertical bars denoting the size of N i (T) are missing at all three occurrences.
11-20:
1. It is stated that |N i (T)| ≤ 2n-3-5 for i ≤ 2, but in a ladder tree is, e.g., |N 1 (T)| = (n-1)(2n-1-5) +
(2n-1-3) > 2n-3-5.
2. Also, it is stated that |N i (T)| > 2n-3-5 for i > 2, but in a ladder tree is, e.g., |N i (T)| = 1 * 2n-
(2i-1)-5 < 2n-3-5 for 2 ≤ i ≤ n-2.
3. Therefore, also the statements on decreasing / increasing the actual values to the bound 2n-
3-5 are the wrong way round. Possibly, this can be fixed by swapping the inequality signs.
4. Is it actually obvious enough that setting all |N i (T)| to 2n-3-5 leads to an upper for the
quotient representing the average moved subtree size? A short explanation might make this
point clearer.
5. In the last sentence of the proof, the weighted average uses a proportion of 1⁄2 for both small
and large subtree (which is okay), but why is the average moved subtree size of small
substrees assumed to be one? Shouldn't the one small subtree with two leaves (which always
exists) make it slightly larger than one?
32: Is there are special reason for (re)defining the mass transport term W 1,n ?
37: The verb 'increase' misses the third-person singular s (as it refers to 'the size of the trees').
45: Only one of the two occurrences of 'thus' is needed.
Page 24
14: Adding a short explanation why the number of unmatched trees is in O(kn) might help the
reader understanding the argument at this point better.
19: The innermost pair of parentheses around 'O(n)' is not necessary.
30: Replacing the word 'a' before 'distance' by 'at' or 'of' conveys the meaning better.
31: The equality sign in the equation might be less fitting than a <, because it is previously (line 25)
stated that S n has a strictly larger neighbourhood.
46-47: Shortly explaining how the bounds for the distance between neighbours of T and S lead to
bounds for W 1 (m T ,m S ) could make it easier for the reader to grasp the idea how the proof works.
Page 25
29: Possibly it should be 'each other' instead of 'each either'.40-41: The actual meaning of the phrase 'a greater understanding of the distribution of tree
neighborhoods with respect to one another' is not very clear (especially due to the underlined part).
Page 26
10: A short remark on how the term '1 + μ' is actually derived would make this passage easier /
faster to comprehend
14: Following Corollary 5.4 it should be 'at least γ' and not 'at most γ' paired trees (and, otherwise,
the argumentation is not sensible at this point).
19: The rationale for bounding the degree by using the maximum possible absolute difference
(irrespective of the actual move) seems a bit unclear. Currently, it does not look like the line of
argument leads to the maximum μ: using the maximum degree of T (instead of the actual,
potentially smaller, value) does not result in the maximum μ because it leads to an unnecessarily
large denominator.
20:
1. The last inequality ('≤ 1⁄2n2-n') holds only if n is even. For odd n, 1⁄2 has to be added to the
upper bound (which is then also okay for the even case). In the end, this change also affects
the curvature bound.
2. It looks like the factors 'k' and 'a' are swapped after inserting their values. While this would
not be wrong, it is a potential source of confusion.
39-40: Providing a proof or at least a clear rationale for the statement 'the pair of trees used in the
proof of Lemma 6.5 will always have negative curvature, for all n ≥ 7' would lend more weight to
it.
Page 27
35: The upper bound does not seem to follow as analogously to the lower bound as it is stated.
Using the inequality stated in line 32 and proceeding analogously to the lower bound leads to
κ p (T , S )
2
⩽ ...⩽ κ (T , S)+
p
p∗max (∣ N (T )∣ , ∣ N (S )∣)
where the factor p in the denominator causes a problem, because simply dropping it (in order to
arrive at the bound from the lemma) bounds the value in the wrong direction (as it is at most 1).
Page 28
11-12: According to Lemma 5.5, the quotient of degrees for two adjacent trees is at least 5/6 only
for n ≥ 4, but there is no such assumption in Lemma 6.9 or its proof.
14-18: Using W 1 (T,S) (resp. W' 1 (T,S)) is slightly imprecise as T and S are the trees and not the
probability densities.
26: Within the min-term of the last case, w is the argument of all four probability densities.
Probably, two of them should have z as their argument.29-33: Using W 1 (resp. W' 1 ) without the arguments for the values of the Wasserstein distance is
slightly imprecise.
50: The word 'the' seems to be missing before 'understanding'.
55: The statement 'rSPR graph degree changes depend quadratically on the product of the size of the
regrafted subtree with its change in depth' seems a bit mixed up, because (according to the proven
lemma) the changes depend linearly on the product (which in turn is quadratic in the number of
leaves).
Page 29
13-17: The same problem as on page 17, lines 15-23.
29: Changing 'achieve' to 'to be achieved' seems a bit more readable.